---
title: "Capstone"
author: "AJ Thurston"
date: "2023-07-14"
format: html
editor: visual
link-citations: true
bibliography: references.bib
---

## Setup

When you click the **Render** button a document will be generated that includes both content and the output of embedded code. You can embed code like this:

```{r setup}
library(summarytools)
library(tidyverse)
library(foreign)
library(lavaan)
library(openxlsx)
library(gtools)
library(lm.beta)
library(domir)
```

## Data Codebook
reference	jobsat	turnover
`refer`: Employee was referred by current employee\
0. Not referred\
1. Referred\
-9. Decline to respond

`jobsat`: Employee job satisfaction: Overall, I am satisfied with my job\
1. Strongly disagree\
2. Disagree\
3. Neither agree nor disagree\
4. Agree\
5. Strongly agree\

`turnover`: Whether or not the employee voluntarily turned over\
0. Active employee\
1. Voluntarily turnover\

## Data Parameters

```{r parameters}
cor <- read.xlsx("cormat.xlsx", sheet = "cor", colNames = T) 
cor[upper.tri(cor)] <- t(cor)[upper.tri(cor)] 
cor <- as.matrix(cor)

sd <- read.xlsx("cormat.xlsx", sheet = "sd", colNames = T) %>% 
  as.matrix() %>% 
  unlist() %>%
  array() %>% 
  
  as.numeric()
mu <- read.xlsx("cormat.xlsx", sheet = "mu", colNames = T) %>% as.matrix()%>%
  unlist() %>%
  array() %>%
  as.numeric()

cov <- cor2cov(cor,sd)
cov <- cov + diag(ncol(cov))*0.1
```

## Data Generation

```{r generation}
set.seed("01234")
df_raw <- MASS::mvrnorm(n = 538, mu = mu, Sigma = cov, tol = .2) %>%
  as.data.frame()
colnames(df_raw) <- colnames(cor)
```

## Data Recode

```{r recode}
df <- df_raw %>%
  mutate(refer_f = (quantcut(refer,   q = c(0,mean(refer)+.05,1), labels = F)-2)*-1) %>%
  mutate(jobsat_skewed = log10(max(jobsat+1) - jobsat)) %>%
  mutate(jobsat_f= (cut(jobsat_skewed,  breaks = 5, labels = F))) %>%
  mutate(turnover_f = (quantcut(turnover,   q = c(0,mean(turnover),1), labels = F)-2)*-1) %>%
  select(refer_f,jobsat_f,turnover_f)

freq(df)
colnames(df) <- c("refer","jobsat","turnover")
cor(df)

write.csv(df, "turnover.csv", row.names = FALSE) 
```